---
title: "ITS2_diversity"
author: "Bailie Wynbelt"
date: "2023-03-28"
output: html_document
---

#install tidyverse

```{r install packages}
library(tidyverse)
```

#load in ITS2 data

```{r read.csv}
reads_WTU_ITS2 <- read_csv("../data/updated_ITS2_reads_WeeTU.csv")
```

###Removing Panicum mili and hirt###

#Remove any rows that have a WTU species value of 122(mili) and 121(hirt)

```{r}
panicum_species <- reads_WTU_ITS2 %>% 
  select(Genus, Species, WTU.species) %>%
  filter(Genus == "Panicum") %>% 
  distinct()

pani_mili_count <- panicum_species %>% 
  filter(Species == "Panicum miliaceum")

pani_hirt_count <- panicum_species %>% 
  filter(Species == "Panicum hirticaule")

no_millet_ITS2 <- reads_WTU_ITS2 %>% 
  filter(WTU.species != "122",
         WTU.species != "121")

```

#Remove data frames that are not needed
```{r}
rm(pani_hirt_count, pani_mili_count, panicum_species)
```

###Remove reads###
```{r}
under_10_reads_ITS2 <- no_millet_ITS2 %>% 
  filter(Reads >= 10)

under_25_reads_ITS2 <- no_millet_ITS2 %>% 
  filter(Reads >= 25)

under_50_reads_ITS2 <- no_millet_ITS2 %>% 
  filter(Reads >= 50)
```

###Diet Diversity###

#Find the sum of total reads
```{r}
under_10_reads_ITS2 <- under_10_reads_ITS2 %>% 
  group_by(SampleID) %>% 
  mutate(sum_reads_ID = sum(Reads))

under_25_reads_ITS2 <- under_25_reads_ITS2 %>% 
  group_by(SampleID) %>% 
  mutate(sum_reads_ID = sum(Reads))

under_50_reads_ITS2 <- under_50_reads_ITS2 %>% 
  group_by(SampleID) %>% 
  mutate(sum_reads_ID = sum(Reads))
```

#Find the proportion of species 
```{r}
under_10_reads_ITS2 <- under_10_reads_ITS2 %>% 
  group_by(SampleID) %>% 
  mutate(pI = (Reads/sum_reads_ID))

under_25_reads_ITS2 <- under_25_reads_ITS2 %>% 
  group_by(SampleID) %>% 
  mutate(pI = (Reads/sum_reads_ID))

under_50_reads_ITS2 <- under_50_reads_ITS2 %>% 
  group_by(SampleID) %>% 
  mutate(pI = (Reads/sum_reads_ID))
```

#Calculate the diversity index
```{r}
under_10_reads_ITS2 <- under_10_reads_ITS2 %>% 
  mutate(diversity_index = -sum(pI*log(pI))) 

under_25_reads_ITS2 <- under_25_reads_ITS2 %>% 
  mutate(diversity_index = -sum(pI*log(pI)))

under_50_reads_ITS2 <- under_50_reads_ITS2 %>% 
  mutate(diversity_index = -sum(pI*log(pI)))

```

#group by species
```{r}
diversity_under_10 <- under_10_reads_ITS2 %>% 
  select(SampleID, sum_reads_ID, pI, diversity_index) %>% 
  distinct(SampleID, .keep_all = TRUE)

diversity_under_25 <- under_25_reads_ITS2 %>% 
  select(SampleID, sum_reads_ID, pI, diversity_index) %>% 
  distinct(SampleID, .keep_all = TRUE)

diversity_under_50 <- under_50_reads_ITS2 %>% 
  select(SampleID, sum_reads_ID, pI, diversity_index) %>% 
  distinct(SampleID, .keep_all = TRUE)
```

### Create synthesized csv's ###

#Join cutoff data frames
```{r}
joined_diversity <- diversity_under_10 %>% 
  left_join(diversity_under_25, by = "SampleID") %>% 
  rename(diversity_index_10 = diversity_index.x,
         diversity_index_25 = diversity_index.y) %>% 
  select(-c(sum_reads_ID.x, pI.x, sum_reads_ID.y, pI.y))

joined_diversity <- joined_diversity %>% 
  left_join(diversity_under_50, by = "SampleID") %>% 
  rename(diversity_index_50 = diversity_index) %>% 
  select(-c(pI, sum_reads_ID))

write_csv(joined_diversity, file = "../outputs/diversity_by_species_ITS2.csv")

```

#Join cutoff data frames 2.0
```{r}
#help im stuck
```

### Visualize/understand ###

#Read_csv
```{r}
diversity <- read_csv("../outputs/diversity_by_species_ITS2.csv")
```

#Look at differences in cut-offs
```{r}
sum_diversity_species_ITS2 <- diversity %>% 
  summarise(mean_diversity_10 = mean(diversity_index_10),
            mean_diversity_25 = mean(diversity_index_25, na.rm = TRUE),
            mean_diversity_50 = mean(diversity_index_50, na.rm = TRUE),
            sd_diversity_10 = sd(diversity_index_10),
            sd_diversity_25 = sd(diversity_index_25, na.rm = TRUE),
            sd_diversity_50 = sd(diversity_index_50, na.rm = TRUE))

```

#Look at ggplot
```{r}
ggplot(data = joined_diversity, aes(x = diversity_index, colors = cutoff)) +
  geom_histogram()
```

#Run stats - is there a statistical differenece
```{r}
#what is the best to run??
diversity_anova <- aov(diversity_index ~ cutoff, data = diversity)
summary(signal_anova)
```